% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %             Discussion                %
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}\label{sec:discussion}
Our results demonstrate that \acgs{} provides a robust and practical solution to the evolutionary governance gap. The framework's core innovations---co-evolutionary adaptation, automated policy synthesis, and democratic oversight---collectively enable a new form of intrinsic, adaptive AI governance.

\subsection{Theoretical and Practical Implications}
Theoretically, our work provides the first formal model of co-evolutionary governance with proven stability guarantees. The successful empirical validation of the Lipschitz constant and convergence rate bridges the gap between AI governance theory and practice. Practically, \acgs{} offers a concrete architectural blueprint for building trustworthy AI systems. By embedding governance directly into the operational loop, the framework moves beyond reactive, external oversight to proactive, ``compliance-by-design'' enforcement. The production deployment on Solana validates its applicability to high-stakes, decentralized environments~\cite{solana2020, quantumagi2024}.

\subsection{Limitations and Sociotechnical Challenges}
While our results demonstrate the first technically viable framework for constitutional evolution at machine speed, several limitations require ongoing research and validation.

\textbf{Distance Metric Validation:} Our Lipschitz stability analysis relies on semantic and behavioral distance metrics that, while empirically validated across multiple embedding models (achieving $L < 1$ across all tested configurations), require continued refinement as constitutional complexity increases. The sensitivity analysis demonstrates robustness across parameter variations, but long-term stability under adversarial constitutional amendments remains an active research area.

\textbf{Democratic Mechanism Scalability:} Our Constitutional Council design addresses documented failure modes from participatory AI initiatives like Sidewalk Labs, implementing anti-capture mechanisms and binding participation protocols grounded in Habermasian procedural legitimacy. However, scaling deliberative democracy to global AI governance while maintaining democratic legitimacy presents ongoing challenges that extend beyond technical solutions.

\textbf{LLM Reliability and Formal Verification Integration:} While our multi-tier validation pipeline achieves 99.7\percent{} policy accuracy after validation, the underlying reliability of LLMs for generating nuanced policy logic remains a research frontier. Our integration of SMT solvers with LLM-based synthesis represents a novel approach, but the completeness of formal verification for complex constitutional principles requires continued development.

\textbf{Adversarial Constitutional Evolution:} Our system detected 93.8\percent{} of adversarial attempts targeting individual policies, but sophisticated attacks on the constitutional evolution process itself---such as coordinated amendment campaigns designed to gradually weaken governance---require further research into constitutional immune systems and meta-governance protections.
