%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  MAIN DOCUMENT: AlphaEvolve-ACGS Research Paper                               %
%  Modular structure with separate files for each section                       %
%  DATE: June 13, 2025                                                         %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Include preamble with packages and custom commands
\input{preamble}

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %          Document Starts Here         %
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\title{\textbf{\acgsbold{}: A Co-Evolutionary Framework for LLM-Driven Constitutional Governance in Evolutionary Computation}}

\author{
    \textbf{Martin Honglin Lyu}\\
    \textit{Soln AI}\\
    Toronto, Ontario, Canada\\
    \texttt{martin@soln.ai}
}

\date{June 13, 2025}

\maketitle

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %               Abstract                %
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Evolutionary computation (EC) systems exhibit emergent behaviors that static governance frameworks cannot adequately control, creating a critical gap in AI safety and alignment. We introduce \acgs{}, a co-evolutionary constitutional governance framework that dynamically adapts alongside evolving AI systems. Our approach integrates four key innovations: (1) a hierarchical architecture that translates high-level constitutional principles into verifiable runtime constraints; (2) an LLM-driven synthesis engine that generates executable policies from natural language; (3) a real-time Prompt Governance Compiler (PGC) for efficient enforcement; and (4) a democratic oversight model with cryptographically-secured amendment and appeal processes.

Empirical validation through the \quantumagi{} production deployment on the Solana blockchain demonstrates the framework's effectiveness. We achieve \textbf{94.7\percent{} constitutional compliance}, a significant improvement from an ungoverned baseline of 31.7\percent{}, while maintaining evolutionary performance. The system operates with policy violation rates reduced to \textbf{<0.5\percent{}} while achieving \textbf{42.3\ms{}} average enforcement latency for policy validation operations. The ACGS-1 production system demonstrates \textbf{85.7\percent{} service availability} with 6 of 7 core services operational, while addressing dependency integration challenges. The measured Lipschitz constant of $\lipschitz \approx 0.74$ and convergence in~14 iterations empirically validate our theoretical stability guarantees. \acgs{} contributes a production-validated methodology for embedding dynamic, verifiable, and constitutionally-grounded governance within~AI, establishing a new paradigm for trustworthy autonomous systems where governance is intrinsic rather than external.
\end{abstract}

\textbf{Keywords:} AI Governance, Evolutionary Computation, Constitutional AI, Large Language Models, Policy-as-Code, Open Policy Agent, Responsible AI, Algorithmic Governance, Dynamic Policy.

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %            Include Sections           %
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Introduction
\input{01-introduction}

% Related Work
\input{02-related-work}

% Methods
\input{03-methods}

% Results
\input{04-results}

% Discussion
\input{05-discussion}

% Future Work
\input{06-future-work}

% Enhanced Constitutional Analyzer
\input{07-enhanced-constitutional-analyzer}

% Quantumagi
\input{08-quantumagi}

% Conclusion
\input{09-conclusion}

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %           Acknowledgments             %
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgments}
We gratefully acknowledge our Constitutional Council simulation participants and domain experts whose insights shaped the AC principles. We thank the open-source maintainers of OPA, vLLM, and Kafka for their guidance, and funding from Google Cloud, AWS, Azure, and NSF that enabled our large-scale evaluation. The authors assume sole responsibility for any remaining errors.

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %             Bibliography              %
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{\small
\raggedright
\bibliographystyle{plainnat}
\bibliography{references}
}

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %               Appendix                %
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{08-appendix}

\end{document}
