\documentclass[11pt, letterpaper, logo, copyright]{googledeepmind}


% \usepackage[final]{neurips_2025}


\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{xcolor}

\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors
\usepackage{booktabs}
\usepackage{pgfplotstable}
\usepackage{amsmath} % For \DeclareMathOperator if needed, or just use text
\usepackage{etoolbox} % For \csdef, robust macro definitions
\usepackage{listings}
\usepackage{xcolor}

\usepackage{siunitx}       % For aligning numbers by decimal point and units
\usepackage{multirow}      % For multi-row cells (e.g., for N spanning header rows)
\usepackage[authoryear, sort&compress, round]{natbib}
\bibliographystyle{abbrvnat}
\usepackage{booktabs}
\usepackage[font=small]{caption}
\usepackage{makecell}      % For line breaks within table cells (useful for long headers)


\lstdefinelanguage{json}{
    basicstyle=\ttfamily\small,
    numbers=left,
    numberstyle=\tiny\color{gray},
    stepnumber=1,
    numbersep=5pt,
    showstringspaces=false,
    breaklines=true,
    frame=single,
    backgroundcolor=\color{gray!5},
    literate=
     *{0}{{\textcolor{blue}{0}}}{1}
      {1}{{\textcolor{blue}{1}}}{1}
      {2}{{\textcolor{blue}{2}}}{1}
      {3}{{\textcolor{blue}{3}}}{1}
      {4}{{\textcolor{blue}{4}}}{1}
      {5}{{\textcolor{blue}{5}}}{1}
      {6}{{\textcolor{blue}{6}}}{1}
      {7}{{\textcolor{blue}{7}}}{1}
      {8}{{\textcolor{blue}{8}}}{1}
      {9}{{\textcolor{blue}{9}}}{1}
      {:}{{\textcolor{red}{:}}}{1}
      {,}{{\textcolor{red}{,}}}{1}
      {"}{{\textcolor{black}{"}}}{1},
}

\usepackage{float}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage{float}
\usepackage{fontawesome5}
\definecolor{darkred}{RGB}{139,0,0} % You can tweak these values
\definecolor{azure}{RGB}{0,127,255}      % define “azure” (adjust as you like)
\usepackage{tcolorbox}
\tcbset{
  azurebox/.style={
    colback=white,                       % box background
    colframe=azure,                      % box frame color
    sharp corners,                       % square corners
    boxrule=1pt,                         % thickness of frame
    left=4pt, right=4pt, top=4pt, bottom=4pt
  }
}
\usepackage{wrapfig}
\usepackage{amsmath}   % For mathematical symbols
\usepackage{amsfonts}  % For mathematical fonts
\usepackage{algorithmic} % For the algorithmic environment
\newcounter{customalgo}
\renewcommand{\thecustomalgo}{\arabic{customalgo}} % Numbering style for the algorithm
\newcommand{\customalgocaption}[1]{%
  \refstepcounter{customalgo}% Increments counter and allows \label to refer to it
  \textbf{Algorithm \thecustomalgo} #1% Formats the caption
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% USER DEFINED
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{math_commands}
\input{commands}
\input{style_files}

\title{Robust Reward Modeling via Causal Rubrics}

\correspondingauthor{{\{pragyahere, hrman, rahulmadhavan, araghuveer, karthikeyanvs, doinap\}@google.com}}

\author[1*]{Pragya Srivastava}
\author[1*]{Harman Singh}
\author[1*]{Rahul Madhavan}
\author[2,3]{\\Gandharv Patil}
\author[1]{Sravanti Addepalli}
\author[1]{Arun Suggala}
\author[1]{Rengarajan Aravamudhan}
\author[1]{Soumya Sharma}
\author[1]{Anirban Laha}
\author[1]{\\Aravindan Raghuveer}
\author[1]{Karthikeyan Shanmugam}
\author[1,3]{Doina Precup}

\affil[1]{Google DeepMind}
\affil[2]{McGill University}
\affil[3]{MILA - Quebec AI Institute}
\affil[*]{Equal Contribution}



\input{includes/abstract}

\begin{document}

\maketitle

\input{includes/introduction}
\input{includes/our_contributions}
\input{includes/related_work}
\input{includes/notations_preliminaries}
\input{includes/methodology}
\input{includes/theory}
\input{includes/experiments}
\input{includes/discussion}
\input{includes/ack}



% \clearpage
\bibliography{References}

% APPENDIX
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \newpage
\appendix
\onecolumn

\vspace{1cm}
\hrule
\par\vspace{0.5cm}
{\Large\bfseries\centering 
{Supplementary Material}
\par\vspace{0.5cm}}
\hrule
\vspace{0.5cm}
\noindent These supplementary materials provide additional details, derivations, and experimental results for our paper. The appendix is organized as follows:


\begin{itemize}[leftmargin=1em]
    \item Section \ref{sec:limitations_future_work} discusses potential limitations of this work.

    \item Section \ref{sec:extended_related_works} provides a broader overview of recent related literature. This is an expanded version of the literature covered in the main paper.
    
    \item Section \ref{sec:additional_results} provides some additional set of results. This is an expanded version of the results covered in the main paper.
    
    \item Section \ref{app:rewordbench_creation} provides the detailed steps we took to reproduce the reWordBench benchmark, as proposed in \citet{wu2025rewordbench}.

    \item Section \ref{sec:experimental_details} provides a detailed overview of our experimental setup.

    \item Section \ref{sec:causal_model_details} provides a detailed walk through of how our causal model extends to prior method. We revisit prior works in light of our causal model. It extends on the shorter version provided in Section \ref{sec:preliminaries}.

    \item Section \ref{app:detailed_augmentation_diagrams} provides a walkthrough of the causal details of the core data augmentation strategies.
    
    \item Section \ref{sec:detailed_methodology} provides a detailed walk through of the method used to train the reward model. It extends on the shorter version provided in Section \ref{sec:methodology}.

    \item Section \ref{sec:theoretical_analysis_detailed} provides a detailed analysis of the theory relating to Reward Hacking and how our proposed method mitigates it.
    
    \item Section \ref{sec:qualitative_example_walkthrough} presents a qualitative example of augmented data created from original data using which is used to train \carma{}.
    
    \item Section \ref{sec:prompt_templates} presents a lists of prompt templates that we use to query our models for generating the data.
    
    \item Section \ref{sec:qualitative_examples} presents a qualitative view common failure modes or biases commonly observed in reward models.
  
\end{itemize}
% \vspace{0.5cm}

\input{supplementaries/limitations}
\input{supplementaries/extended_related_works}
\input{supplementaries/additional_results}
\input{supplementaries/rewordbench_reproduce}
\input{supplementaries/experimental_setup_details}
\input{supplementaries/detailed_causal_model}
\input{supplementaries/causal_view_of_augmentation}
\input{supplementaries/detailed_methodology}
\input{supplementaries/theory_detailed}
\input{supplementaries/detailed_example_walkthrough}
\input{supplementaries/prompt_templates}
\input{supplementaries/qualitative_examples_reward_hacking}

\end{document}

