% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %             Discussion                %
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}\label{sec:discussion}
Our results demonstrate that \acgs{} provides a robust and practical solution to the evolutionary governance gap. The framework's core innovations---co-evolutionary adaptation, automated policy synthesis, and democratic oversight---collectively enable a new form of intrinsic, adaptive AI governance.

\subsection{Theoretical and Practical Implications}
Theoretically, our work provides the first formal model of co-evolutionary governance with proven stability guarantees. The successful empirical validation of the Lipschitz constant and convergence rate bridges the gap between AI governance theory and practice. Practically, \acgs{} offers a concrete architectural blueprint for building trustworthy AI systems. By embedding governance directly into the operational loop, the framework moves beyond reactive, external oversight to proactive, ``compliance-by-design'' enforcement. The production deployment on Solana validates its applicability to high-stakes, decentralized environments~\cite{solana2020, quantumagi2024}.

\subsection{Limitations and Sociotechnical Challenges}
Despite promising results, several challenges remain.
\begin{itemize}[leftmargin=*,topsep=2pt,itemsep=2pt,parsep=0pt]
    \item \textbf{LLM Reliability:} While our validation pipeline is effective, the underlying reliability of LLMs for generating nuanced policy logic remains a key research frontier. The risk of semantic misinterpretation or ``loophole'' generation necessitates a continued emphasis on formal verification and human oversight—a challenge that future work might address through AI-assisted human review and enhanced prompt engineering techniques.
    \item \textbf{Constitutional Legitimacy:} The legitimacy of the AC depends on the inclusiveness and fairness of the Constitutional Council. Ensuring diverse representation and preventing ``constitutional capture'' by powerful stakeholders are persistent sociotechnical challenges that require ongoing diligence beyond the technical framework—challenges that could be mitigated through transparent governance mechanisms and algorithmic auditing of decision processes.
    \item \textbf{Scalability of Oversight:} While the technical components scale efficiently, the human oversight mechanisms (Council deliberations, HITL reviews, appeals) create a potential bottleneck. Scaling these social processes without sacrificing quality is a significant challenge—one that future research might address through hierarchical governance structures and AI-augmented decision support systems.
    \item \textbf{Adversarial Robustness:} Our system detected 93.8\percent{} of adversarial attempts in testing, but the risk of novel attacks on the GS engine itself requires further research into more advanced defense-in-depth strategies—an area where multi-model consensus and adversarial training could provide enhanced protection.
\end{itemize}
